
---
title: MALTEM costs overview
author: 
  - Laia Cirera
  - Joe Brew
  - Elisa Sicuri 
fig_height: 2.6
fig_width: 4
header-includes: \usepackage{longtable}
highlight: zenburn
latex_engine: xelatex
output:
  pdf_document:
    includes:
      in_header: mystyle.sty
  html_document: default
tables: yes
toc: yes
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE }
# No scientific notation
options(scipen=999)

#####
# PACKAGES
#####
library(dplyr)
library(xtable)
library(ggplot2)
library(knitr)
library(png)
library(grid)
library(extrafont)
library(tidyr)
library(gridExtra)
library(maptools)
library(rgeos)
library(ggmap)

loadfonts(device = 'postscript', quiet = TRUE)

# BASIC KNITR OPTIONS
opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, fig.height = 3, fig.align='center')
options(xtable.comment = FALSE)
```

```{r, results='hide'}
# GET THEME FOR GGPLOTS
source('theme.R')
# READ IN DATA
source('read_in.R')
```

```{r, fig.align='center', fig.height = 3}
# TITLE MAP
ggplot() +
  coord_map() +
  geom_polygon(data = moz3_fortified,
       aes(x = long, y =lat, group = group), fill = 'grey', color = 'white') +
    geom_polygon(data = maputo_fortified,
       aes(x = long, y =lat, group = group), fill = 'darkred', color = NA, alpha = 0.6) +
  theme_tufte() +
  theme(axis.ticks.length = unit(0.001, "mm")) + labs(x=NULL, y=NULL) +
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      # legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())
```

Full code at https://github.com/joebrew/maltem_costs

\newpage


# Introduction  

What falls is a brief overview of costs associated with the MALTEM program.

## By area

```{r, fig.height = 4.5}

# Distribution of all costs by objreal
temp <- df %>%
  group_by(Area = toupper(OBJReal)) %>%
  summarise(n = n(),
            dollars = sum(USD)) %>%
  mutate(p = dollars / sum(dollars) * 100) %>%
  arrange(desc(p)) %>%
  filter(!is.na(Area))
  
# Order factor levels by p
temp$Area <- factor(temp$Area, levels = temp$Area)

ggplot(data = temp,
       aes(x = Area, y = p)) +
  geom_bar(stat = 'identity',
           fill = 'darkred',
           alpha = 0.6) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 10)) +
  xlab('Area') +
  ylab('Percent of total expenditure') +
  geom_text(aes(y = p + 2 ,
                label = paste0(round(p, 2), '%')),
            size = 2) +
  ggtitle('All costs by area (excluding unclassified)')

```

## MDA expenditures over time


```{r}

# MDA expenditures over time
mda_time <- mda %>%
  group_by(Datapago) %>%
  summarise(dollars = sum(USD)) %>%
  mutate(dollars_cum = cumsum(dollars))

# Month specific MDA expenditures over time
mda_month <- mda_time %>%
  mutate(month_numer = format(Datapago, '%m'),
         month = format(Datapago, '%B')) %>%
  group_by(month, month_numer) %>%
  summarise(dollars = sum(dollars)) 
mda_month <- mda_month[order(as.numeric(mda_month$month_numer)),]
mda_month$month <- factor(mda_month$month, levels = mda_month$month)
ggplot(data = mda_month,
       aes(x = month, y = dollars)) +
  geom_bar(stat = 'identity', fill = 'darkred', alpha = 0.6) +
  xlab('Month') +
  ylab('Dollars spent') +
  ggtitle('Month-specific MDA expenditures over time') +
    theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 10))


```


```{r}

ggplot(data = mda_time,
       aes(x = Datapago, y = dollars_cum)) +
  geom_area(fill = 'darkred', alpha = 0.6) +
  xlab('Date') +
  ylab('Dollars spent') +
  ggtitle('Cumulative MDA expenditures over time')

# ADD FUTURE ESTIMATIONS

```


## MDA expenditures by item

```{r, fig.height = 4.5}
# MDA by item
by_item <- 
  mda %>%
  group_by(Item) %>%
  summarise(n = n(),
            dollars = sum(USD)) %>%
  arrange(desc(dollars)) %>%
  mutate(p = dollars / sum(dollars)  * 100)

by_item$Item <- factor(by_item$Item, levels = by_item$Item)

ggplot(data = by_item,
       aes(x = Item, y = p)) +
  geom_bar(stat = 'identity', 
           fill = 'darkred',
           alpha = 0.6) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 10)) +
  ylab('Percent') +
      geom_text(aes(y = p + 2 ,
                label = paste0( round(p, 2), '%')),
            size = 2) 


```

# IRS details


```{r}
# Get cmulative counts
for (j in 2:ncol(irs)){
  column_name <- names(irs)[j]
  new_name <- paste0(column_name, '_p')
  irs[,new_name] <- cumsum(irs[,j])
}
```


## Houses sprayed

```{r, fig.height = 8}
g1 <- ggplot(data = irs,
       aes(x = week,
       y = houses_sprayed)) +
  geom_bar(stat = 'identity',
           fill = 'darkgreen',
           alpha = 0.6) +
  xlab('Week') +
  ylab('Houses sprayed') +
  ggtitle('Houses sprayed by week')

g2 <- ggplot(data = irs,
       aes(x = week,
       y = houses_sprayed_p)) +
  geom_area(stat = 'identity',
           fill = 'darkgreen',
           alpha = 0.6) +
  xlab('Week') +
  ylab('Houses sprayed') +
  ggtitle('Cumulative houses sprayed by week')

multiplot(g1, g2)


```

## Structures sprayed

```{r, fig.height = 8}

g1 <- ggplot(data = irs,
       aes(x = week,
       y = n_structures)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  xlab('Week') +
  ylab('Structures sprayed') +
  ggtitle('Structures sprayed by week')

g2 <- ggplot(data = irs,
       aes(x = week,
       y = n_structures_p)) +
  geom_area(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  xlab('Week') +
  ylab('Structures sprayed') +
  ggtitle('Cumulative structures sprayed by week')

multiplot(g1, g2)
```

## People protected  


```{r, fig.height = 8}

g1 <- ggplot(data = irs,
       aes(x = week,
       y = n_protected)) +
  geom_bar(stat = 'identity',
           fill = 'darkblue',
           alpha = 0.6) +
  xlab('Week') +
  ylab('People protected') +
  ggtitle('People protected by week')

g2 <- ggplot(data = irs,
       aes(x = week,
       y = n_protected_p)) +
  geom_area(stat = 'identity',
           fill = 'darkblue',
           alpha = 0.6) +
  xlab('Week') +
  ylab('People protected') +
  ggtitle('Cumulative people protected by week')

multiplot(g1, g2)
```

\newpage

## IRS costs  

(Simply dividing by the total direct cost of `r total_direct_costs`.)

```{r}
temp <- data.frame(
  Cost = 
   c(total_direct_costs /
    sum(irs$houses_sprayed),
   total_direct_costs / 
    sum(irs$n_structures),
  total_direct_costs / 
    sum(irs$n_protected)),
  Category = c('Per house',
               'Per structure',
               'Per person')) %>%
  arrange(Cost)

temp$Category <- 
  factor(temp$Category,
         levels = temp$Category)

```

```{r}

ggplot(data = temp,
       aes(x = Category,
           y = Cost)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           color = 'black',
           alpha = 0.4) +
  ggtitle('Cost per outcome') +
    geom_text(aes(y = Cost + 2 ,
                label = paste0('$', round(Cost, 2))),
            size = 5) 

```


## Expenses over time 

(Direct costs only, assuming fixed house cost of $`r round(temp$Cost[temp$Category == 'Per house'], digits = 2)`.)

```{r}
irs$cost <- irs$houses_sprayed * temp$Cost[temp$Category == 'Per house']

 ggplot(data = irs,
       aes(x = week,
       y = cost)) +
  geom_area(stat = 'identity',
           fill = 'darkorange',
           color = 'black',
           alpha = 0.6) +
  xlab('Week') +
  ylab('Dollars') +
  ggtitle('Cumulative costs per week')

```